### 2013-03-26 doc/final_report/design.tex
Started writing the implementation chapter

>>>>

Despite the use of an iterative development methodology, some of the key
components within Partridge were designed up front in order to accelerate
implementation and reduce time overheads when adding new components to the
system. The components that make up Partridge and the ways in which they react
are described and analysed below. Choices regarding the type of application,
the target audience and the programming environment are also discussed.

### 2013-03-26 doc/final_report/implementation.tex
Started writing the implementation chapter

@@

\subsection{ Sourcing and Acquiring Scientific Papers}
For Partridge to be a success, it is important to include a large variety of
different scientific papers in its corpus. Therefore, several sources of
scientific papers were explored and considered.

The most convenient and accessible source is the ART corpus that SAPIENTA was
trained with\cite{citeulike:11077287}. This corpus is already stored using the
CoreSC schema and has been pre-annotated. This means that Partridge would not
need to do any conversion or pre-processing on the papers. These scientific
papers are all chemistry papers. Therefore, more papers from other scientific
domains are needed to get a wider variety of topics and make Partridge useful
to as wide an audience as possible.

The `mega-journals' PLOSOne \url{http://www.plosone.org/} and arXiv
\url{http://www.arxiv.org/} were suggested as sources for more open access
articles that could be added to Partridge. Both of these sites were found to
contain large volumes of open access papers. Most of the journals stored on the
arXiv site were in PDF format which, as discussed below, are difficult to
convert and annotate. However, PLOSOne use the SciXML markup language for
papers published through their journal, which made converting and annotating
them a lot simpler for Patridge. PLOSOne also publish all of their papers under
the Creative Commons Share-Alike-By-Attribution license\cite{ccbyattr}. This
meant that as long as the author information was left intact, all of the papers
could be used for data mining purposes.

\subsection{Format Conversion}
Most scientific papers available on the internet are formatted as PDF
documents. However, Partridge uses and stores documents that use the CoreSC
schema by Soldatova and Liakata\cite{liakata2008guidelines}. Therefore some
spike work was carried out to determine the feasibility of converting papers
published as PDF documents into XML documents. Townsendi \emph{et al.} (2009) liken converting
PDF to XML to ``converting hamburgers into cows," they go on to explain that
PDF documents do not contain any semantic data and documents lose much of their
explicit structure when they are formatted in this way \cite{Townsend2009}.
Therefore, to convert PDF documents into an NLP-friendly format, some
heuristics must be used to detect the document's structure\cite{pdfminer}.

@@

Once all the data files and libraries were successfully in place, the Python
version of SAPIENTA was used to process some of the papers converted from PDF.
This was a great success and the annotations provided by SAPIENTA seemed to be
accurate.

@@

@@

@@

>>>>

Implementing Partridge from a set of initial designs proved to be a
challenging but rewarding process. There were a great number of technical
challenges involved in building the system, a large proportion of the
implementation time was spent working on the paper preprocessor module.
However, some aspects of the web interface and backend were also quite
difficult to tackle.

@@

\subsection{ Sourcing and Acquiring Scientific Papers}

In order to provide a useful service to researchers looking for papers, growing
a large corpus of scientific literature was made a high priority during the
project. This was especially important when training the Machine Learning
models since training data sets need to be statistically significant in order
to provide any meaninful results. Users are given the option, and actively
encouraged, to upload papers that they own or publish on another author's
behalf to the Partridge instance. However, to speed up this process, a number
of papers were acquired from various open access sources.

The most convenient and accessible source was the ART corpus that SAPIENTA was
trained with\cite{citeulike:11077287}. This corpus was already stored using the
CoreSC schema and had been pre-annotated. This meant that Partridge would not
need to do any conversion or pre-processing on the papers. The papers in this
corpus were all of a similar type and covered similar subject areas within the
domain of Biochemistry. There were also approximately 260 papers in the
collection. These papers were used to help establish a conversion process from
PDF to annotated XML. However, they only formed part of the final corpus.
Including a large number of papers from other sources provided a more
comprehensive collection of data for Partridge to learn from.

The `mega-journals' PLOSOne \url{http://www.plosone.org/} and arXiv
\url{http://www.arxiv.org/} were suggested as sources for more open access
articles that could be added to Partridge. Both of these sites were found to
contain large volumes of open access papers. Most of the articles stored on the
arXiv site were in PDF format which, as discussed below, are difficult to
convert and annotate. However, PLOSOne use the SciXML markup language for
papers published through their journal, which made converting and annotating
them a lot simpler for Patridge. PLOSOne also publish all of their papers under
the Creative Commons Share-Alike-By-Attribution license\cite{ccbyattr}. This
meant that as long as the author information was left intact, all of the papers
could be used for data mining purposes.

A third mega-journal used for downloading papers was the PubMed Centrala (PMC)
repository \url{http://www.ncbi.nlm.nih.gov/pmc/}. Not all of
the papers available for viewing at the PMC website are open access. However,
they do offer a listing for the subset of open access papers available on their
website. The PubMed Central format is also very similar to the SciXML format
and compatible with SAPIENTA. Therefore, a large number of PubMed papers were
also used for the initial training of Partridge.

\subsection{Format Conversion} Most scientific papers available on the internet
are formatted as PDF documents. However, Partridge uses and stores documents as
XML markup and uses the CoreSC schema by Soldatova and
Liakata\cite{liakata2008guidelines} for annotation. Therefore some spike work
was carried out to determine the feasibility of converting papers published as
PDF documents into XML documents. Townsend \emph{et al.} (2009) liken
converting PDF to XML to ``converting hamburgers into cows," they go on to
explain that PDF documents do not contain any semantic data and documents lose
much of their explicit structure when they are formatted in this way
\cite{Townsend2009}.  Therefore, to convert PDF documents into an NLP-friendly
format, some heuristics must be used to detect the document's
structure\cite{pdfminer}.

@@

Once all the data files and libraries were successfully in place, the Python
version of SAPIENTA was used to process some of the papers converted from PDF.
This appeared to have been successful. However, after applying some machine
learning evaluation techniques as discussed in Chapter \ref{chapter:testing},
it was clear that the Python version of SAPIENTA was not accurate enough to
provide CoreSC annotations, scoring 44\% accuracy on average. After discussion
with its author, it became apparent that some of the features provided in the
server-side version of SAPIENTA had not been implemented in the Python code
yet. Therefore, it was necessary to switch bak to remote annotation.

@@

Once a large number of papers had been acquired from ART, PlosOne and PubMed
Central, work could begin on implementing a paper type classifier able to
discriminate between ``Case Study", ``Research" and ``Review" papers. However,
for a machine learning classifier to discriminate between classes of data, it
is necessary to choose a set of features that differ as much as possible
between the classes. For example, to discriminate between cats and dogs,
``number of legs" is a poor feature since both species normally have 4.
However, the noise that these animals make may be a good discriminative feature
since cats tend to mew and dogs bark.

\subsubsection{ Feature Selection and Clustering}

There are several well established features used for classifying text within
the NLP community. However, since Partridge has access to CoreSC annotations
for each paper, an investigation was carried out into how useful CoreSC data
can be as a feature for paper type discrimination. Every sentence in a given
annotated paper has an associated CoreSC label. It is therefore trivial to
calculate the proportion of a paper made up of sentences with a specified
CoreSC type. Several papers were manually inspected and the proportions of each
CoreSC represented within were calculated. A pattern quickly emerged between
the types of paper and the proportions of each CoreSC within them.

\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{images/implementation/review_corescs.png}
\caption{Contrasting CoreSC content of Review and Research Papers}
\label{fig:coresc_pies}
\end{figure}

Figure \ref{fig:coresc_pies} shows the CoreSC content of a review aper and a
research paper randomly selected from the corpus. The review papers tend to be
made up almost entirely from Background CoreSC sentences. However, research
papers are much more evenly spread, made up of several different types of
CoreSC. This investigation suggested that there is almost certainly a discriminative
relationship between CoreSC categories and a paper's type.

Although all of the CoreSC types could be influential in determining a paper's
type, it was likely that some were more important than others. A further
investigation was carried out to discover the most effective features for
categorising the data. Since no classifier model had been trained at this
point, the data was clustered into three unlabelled groups using the K-means
clustering algorithm. This algorithm was run with three CoreSC

@@

During the development of Partridge, the structure of the SQL database

@@

One of the most important user suggestions for improving the usability of
Partridge and facilitating the rapid expansion of the paper corpus was the
Paper Upload Bookmarklet. The bookmarklet was not in the initial system design
and was added as an extra feature towards the end of the project. It allows
users browsing sites such as PubMed Central and PlosOne to add papers that
interest them to Partridge without having to download the paper from the
journal and upload it to the Partridge server manually.

The system is implemented as a standard web page wrapped inside a javascript
bootstrap module that can be added as a bookmark to a user's web browser. When
the user finds a paper that they wish to add to Partridge on one of the
compatible journal sites, they click the bookmark, loading an embedded frame in
their current browser window and automatically uploading the paper that the
user is looking at to Partridge's server.

The bookmarklet trivialises the act of adding new papers to Partridge and this
should help to encourage users to help grow Partridge's paper corpus.

### 2013-03-26 doc/final_report/methodology.tex
Started writing the implementation chapter

\section{Existing Methodologies}
Selecting a suitable development methodology for building Partridge is another
very important choice for the project.

@@

\subsection{ Work Timeline }

The tasks involved in Partridge have been carefully calculated and prioritised.
They were then added to the GitHub issue management system and a report
generated listing them in the order that they are expected to be
accomplished. This report can be seen in Appendix \ref{sec:timeline}.

>>>>

Chosing an effective development methodology can play a very large part in the
success or failure of a software project. A number of existing development
methodologies have been developed and used very effectively over the last 50
years. However, custom methodologies are also commonly used for projects that
aren't suited to one of the commonly used approaches. Some common approaches to
software development and the formulation of a custom methodology for Partridge
are discussed below.

\section{Existing Methodologies}

@@

Programming tasks were added to the GitHub project tracker facility in the
same place as the Wiki was hosted. The `milestones' feature, which allows tasks
to be split into groups that must be completed by an arbitrary date, was used
to plan iterations and could be used to provide precise summaries of how the
project was progressing. In January, the project ran behind by a few weeks.
The time tracker explained that the project was precisely 17 days behind and
highlighted the set of tasks that needed to be completed to bring the project
back up to speed.

\section{ Initial Plan and Project Timeline }

During the early stages of the Partridge project, an initial system design was
drawn up and a set of tasks that covered the system's implementation were
formulated. These tasks were divided up into iterations and added to the
project planner on GitHub. Each task's dependencies were carefully considered
in order to prioritise tasks properly. The initial project plan can be seen in
Appendix \ref{sec:timeline}.
